---
title: >
  ISL - Chapter 2 Labs
author: Federico Marini
output: 
  html_document:
  # BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cosmo
    code_folding: show
    code_download: true
editor_options: 
  chunk_output_type: console

---

# Chapter 2 Lab: Introduction to R

This one is indeed a soft intro!

No worry to make things wrong: TRY!

Got R?

- R can be downloaded from http://cran.r-project.org/

Start...

- an R script (tired!)
- an R Markdown document (wired!)

## Basic commands

Functions are noted as `funcname(input1, input2)` - `input`s are called _arguments_

Creating a vector by `c`oncatenating

```{r}
c(1, 3, 2, 5)
x <- c(1, 3, 2, 5) # use <- for assignment, leave = for parameter values
x

x = c(1,6,2)
y = c(1,4,3)
```

Navigating history:

- up arrow
- ctrl + up (RStudio) while typing part of the command

Help:

- use `?`
- `help(funcname)`

Length (of a vector)

```{r}
length(x)
length(y)
```

Adding things together

```{r}
1 + 4
x + y # if same length
x + 1
x + c(1,2)
```


Listing objects, removing them

```{r}
ls()
rm(x, y)
ls()
```

Create a matrix of numbers

```{r}
?matrix
x <- matrix(data=c(1,2,3,4), nrow=2, ncol=2)
x
x <- matrix(data=c(1,2,3,4), 2, 2)
x
# Recommended: specify arguments!
matrix(c(1,2,3,4),2,2,byrow=TRUE)
```

Squaring, squared root

```{r}
x
sqrt(x)
sqrt(x)[,1]
x^2
```

Random normal variables, with `rnorm`

```{r}
# bonus - when dealing with random quantities!
set.seed(42)

x <- rnorm(50)
# default: mean of 0 and standard deviation of 1
y <- x + rnorm(50,mean=50,sd=.1) 
```

Computing the correlation between two sets of numbers

```{r}
cor(x,y)
```

Mean & Standard deviation

```{r}
mean(x)
sd(x)
mean(y)
sd(y)
var(x)
var(y)

set.seed(3)
y <- rnorm(100)
mean(y)
var(y)
sqrt(var(y))
sd(y)
```


## Graphics

Meet the `plot()` function!

`plot(x,y)`

```{r}
?plot
```


The plot() function is the primary way to plot data in R. For instance, plot(x,y) produces a scatterplot of the numbers in x versus the numbers in y. There are many additional options that can be passed in to the plot() function. For example, passing in the argument xlab will result in a label on the x-axis. To find out more information about the plot() function, type ?plot.

```{r}
set.seed(42)
x <- rnorm(100)
y <- rnorm(100)
plot(x,y)
plot(x,y,
     xlab="this is the x-axis",
     ylab="this is the y-axis",
     main="Plot of X vs Y")
# plot()
```
      
Saving the output:

```
open device
plot...
close device
```

```{r}
pdf("Figure.pdf")
plot(x,y,col="green")
dev.off()
```

Devices: `pdf()`, `jpeg()`, ...

Creating `seq`uences of numbers

```{r}
?seq
seq(0,1)
seq(0,1,length=10)
pi
seq(-pi,pi,length=50)
```

Some more complex plots:

- `contour()` to represent 3d data

```{r}
?contour
?outer

# assignment and printing trick ;)
(x <- seq(-pi,pi,length=50))
y <- x
f <- outer(x,y,function (x,y)cos(y)/(1+x^2))
f
contour(x, y, f)
contour(x,y,f,nlevels=45,add=T)
fa <- (f-t(f))/2
contour(x,y,fa,nlevels=15)
```

- `image()` to produce a color-coded plot whose colors depend on the z value (a.k.a. heatmap)

```{r}
image(x,y,fa)
```


- `persp()` to produce a three-dimensional plot - The arguments theta and phi control the angles at which the plot is viewed.

```{r}
persp(x,y,fa)
persp(x,y,fa,theta=30)
persp(x,y,fa,theta=30,phi=20)
persp(x,y,fa,theta=30,phi=70)
persp(x,y,fa,theta=30,phi=40)
```

## Indexing data

Examining a subset of a dataset `A`

```{r}
A <- matrix(1:16,4,4)
A
A[2,3]
A[c(1,3),c(2,4)]
A[1:3,2:4]
A[1,]
A[1:2,]
A[,1:2]
A[-c(1,3),]
```

`dim`ensions

```{r}
dim(A)
```


## Loading data

```{r}
library(ISLR)
Auto
write.table(Auto, file = "Auto.txt")
read.table(file = "Auto.txt")

?read.csv

# fix(Auto)
View(Auto)
knitr::kable(Auto)
class(Auto)
dim(Auto)
names(Auto)
str(Auto)
na.omit(Auto)
dim(na.omit(Auto))
```


## Additional summaries

Scatterplot of quantitative variables

```{r, error=TRUE}
plot(cylinders , mpg)
```


To refer to a variable, we must type the data set and the variable name joined with a `$` symbol.

DO NOT USE `attach()`

```{r}
plot(Auto$cylinders, Auto$mpg)
plot(Auto$cylinders)

# but this is actually a factor...

Auto$cylinders <- as.factor(Auto$cylinders)
plot(Auto$cylinders, Auto$mpg)
plot(Auto$cylinders)

# customizing the aspect
plot(Auto$cylinders, Auto$mpg, col="red", varwidth=T, xlab="cylinders", ylab = "MPG")
plot(Auto$cylinders, Auto$mpg, col="red", horizontal=T, xlab="cylinders", ylab = "MPG")

```

Histogram

```{r}
hist(Auto$mpg)
hist(Auto$mpg, col = 2)
hist(Auto$mpg, breaks=15)
```

Scatterplot matrix

```{r, fig.width=10, fig.height=10}
pairs(Auto)

# for just a subset
pairs(~mpg + displacement + horsepower + weight, Auto)
```


`identify()`: a useful interactive method for identifying the value for a particular variable for points on a plot. 

```{r}
plot(Auto$horsepower, Auto$mpg)
identify(Auto$horsepower,Auto$mpg,Auto$name)
# click, click, click, and then ESC
```

One of your best friends: `summary()`, to produce a numerical summary of each variable in a particular data set

```{r}
summary(Auto)
summary(Auto$mpg)
summary(Auto$cylinders)
```

Done for today

```{r eval=FALSE}
q()
quit()
```

but first, you can `savehistory("history_filename.R")` - still, I would recommend to always use a script anyway!

# Session Info {-}

```{r}
sessionInfo()
```
