---
output: github_document
---

# Home

Welcome to the MachineLearningClub wiki!

## Details

Each meeting will be focused on a particular chapter, or a part thereof. 



### Deep Learning with R

```{r, echo = FALSE}
schedule_df <- data.frame(
  Sections = "",
  Date = "2020-11-XX",
  Presenter = "Max Mustermann",
  Exercises = "",
  stringsAsFactors = FALSE
)
```

```{r, echo = FALSE}
schedule_df <- rbind(
  schedule_df,
  c("Chapter 2", "2020-MM-DD", ""),
  c("Chapter 3", "2020-MM-DD", ""),
  c("Chapter 4", "2020-MM-DD", ""),
  c("Chapter 5", "2021-MM-DD", ""),
  c("Chapter 6", "2021-MM-DD", ""),
  c("Chapter 7", "2021-MM-DD", ""),
  c("Chapter 8", "2021-MM-DD", ""),
  c("Chapter 9", "2021-MM-DD", "")
)  
  # c("Topic2", "2020-03-09", "Person2", ""),
  
```


```{r, echo = FALSE}
knitr::kable(schedule_df)
```


Please use `wiki_schedule.Rmd` in the repo to generate the content for this page


Then, take the content of the generated `wiki_schedule.md` file and replace the content of the Wiki at https://github.com/imbeimainz/MachineLearningClub/wiki


---

### Introduction to Statistical Learning

The first part of each meeting will be dedicated to a short recap on the "Labs" from the previous chapter

```{r, echo = FALSE}
schedule_df <- data.frame(
  Sections = "Intro",
  Date = "2020-03-02",
  Presenter = "Federico",
  Exercises = "",
  stringsAsFactors = FALSE
)
```

```{r, echo = FALSE}
schedule_df <- rbind(
  schedule_df,
  c("Chapter 2", "2020-03-09", "Irene + Federico (labs)"),
  c("rescheduled!", "2020-03-16", ""),
  c("Chapter 3", "2020-03-23", "Philipp + Jan"),
  c("Chapter 4", "2020-03-30", "Jochem + Robert"),
  c("Chapter 5", "2020-04-20", "Dan + Jennifer"),
  c("Chapter 6", "2020-04-27", "Gregor + Rex (Part 1, Gregor)"),
  c("Chapter 10", "2020-05-04", "Toni + Alicia (Part 1, Toni)"),
  c("cancelled", "2020-05-11", ""),
  c("Chapter 6", "2020-05-18", "Gregor + Rex (Part 2, Rex)"),
  c("Chapter 6", "2020-05-25", "Labs, Chapter 6 (Gregor + Rex)"),
  c("holiday break", "2020-06-01", ""),
  c("Chapter 7", "2020-06-08", "Jochem + Irene (Part 1)"),
  c("Chapter 7", "2020-06-15", "Jochem + Irene (Part 2)"),
  c("Chapter 8", "2020-06-22", "Franziska"),
  c("Chapter 9", "2020-06-29", "Manuel + Lars"),
  c("Chapter 10", "2020-07-06", "Toni + Alicia (Part 2, Alicia) + Lab"),
  c("Chapter 9", "2020-07-13", "Lab, Manuel + Recap & wrap up"))
  
  # c("Topic2", "2020-03-09", "Person2", ""),
  
```


```{r, echo = FALSE}
knitr::kable(schedule_df)
```


