---
title: "Chapter 3 Linear Regression"
subtitle: ""
author: "Jan Linke </br> Philipp Mildenberger</br>"
date: "2020/03/16</br>
  IMBEI - University Medical Center Mainz"
output: 
  xaringan::moon_reader:
    css: ["default", "default-fonts","css/animate.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: [center, middle]
---

```{r setup, include=FALSE}
library(rgl)
library(knitr)

options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  eval = TRUE,
  message = FALSE,
  echo = FALSE,
  warnings = FALSE,
  fig.align = "center",
  warning = FALSE,
  error = FALSE
)
opts_knit$set(root.dir = "..") ## knit in parent folder for the sake of easy data access

## fuer 3d graphik
knitr::knit_hooks$set(webgl=hook_webgl)
```

# Chapter Overview

- Motivating Example
- Simple Linear Regression
- Multiple Linear Regression
- Other Considerations in the Regression Model
- Example expained
- Comparison to $K$-Nearest Neighbors

---

# Motivating Example
## Data 
```{r, echo=FALSE}
## laden, relativer Pfad, sollte funktionieren wenn als R-project geoeffnet.
## quick'n'dirty hack: option 1 fuer arbeit in 
try(Advertising <- read.csv("Introduction to Statistical Learning - 2020/chapter3-linear_regression/data/Advertising.csv"),TRUE)
try(Advertising <- read.csv("data/Advertising.csv"),TRUE) 
```

```{r}
head(Advertising)
```

---
class: center, middle

## Questions


* Relationship between advertising budget and `sales`?
* How strong is relationship between advertising budget and `sales`?
* Which media contribute to `sales`?
* How accurately (and precisely) can we estimate the effect of each medium?
* How accurately can we predict future sales?
* Is the relationship linear?
* Is there synergy (i.e. interaction) among the adventising media?

---

# Simple Linear Regression
Predict quantitative response $Y$ on the basis of $X$

$$ Y \approx \beta_0 + \beta_1 X $$ 

$\beta_0$ and $\beta_1$ are unknown constants

---
## Advertising Example 
```{r}
plot(sales ~ TV, data=Advertising, pch=16, col=2)
```

---
## Advertising Example
```{r}
lm1 <- lm(sales~TV, data=Advertising)
coefs <- lm1$coefficients
plot(sales ~ TV, data=Advertising, pch=16, col=2)
abline(a=coefs[1],b=coefs[2], lwd=2, col=4)
```

---
## Advertising Example
```{r}
pred  <- predict(lm1)
plot(sales ~ TV, data=Advertising, pch=16, col=2)
abline(a=coefs[1],b=coefs[2], lwd=2, col=4)
with(Advertising,segments(TV, sales, TV, pred, lwd = 1))
```


# Multiple Linear Regression

```{r, webgl=TRUE}
x <- sort(rnorm(1000))
y <- rnorm(1000)
z <- rnorm(1000) + atan2(x,y)

plot3d(x, y, z, col=rainbow(1000))

```


---

# Other Considerations

---

# Example Explained

---

class: center, middle

# Done!

