---
title: "Chapter 3 Linear Regression"
subtitle: ""
author: "Jan Linke </br> Philipp Mildenberger</br>"
date: "2020/03/16</br>
  IMBEI - University Medical Center Mainz"
output: 
  xaringan::moon_reader:
    css: ["default", "default-fonts","css/FMstyles.css","css/animate.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: [center, middle]
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  eval = TRUE,
  message = FALSE,
  echo = TRUE,
  warnings = FALSE,
  fig.align = "center",
  warning = FALSE,
  error = FALSE
)
```

# Chapter Overview

- Motivating Example
- Simple Linear Regression
- Multiple Linear Regression
- Other Considerations in the Regression Model
- Example expained
- Comparison to $K$-Nearest Neighbors

---

# Motivating Example
## Data 
```{r, echo=FALSE}
## laden, relativer Pfad, sollte funktionieren wenn als R-project geoeffnet.
## quick'n'dirty hack: option 1 fuer arbeit in 
try(Advertising <- read.csv("Introduction to Statistical Learning - 2020/chapter3-LinReg/data/Advertising.csv"),TRUE) 
try(Advertising <- read.csv("data/Advertising.csv"),TRUE) 
```

```{r}
head(Advertising)
```

--

## Questions
* Relationship between advertising budget and `sales`?

--

* How strong is relationship between advertising budget and `sales`?

--

* Which media contribute to `sales`?

---

## Questions (cont'd)

* How accurately (and precisely) can we estimate the effect of each medium?
* How accurately can we predict future sales?
* Is the relationship linear?
* Is there synergy (i.e. interaction) among the adventising media?

---

# Simple Linear Regression
Predict quantitative response $Y$ on the basis of $X$

---

# Multiple Linear Regression

---

# Other Considerations

---

# Example Explained

---

class: center, middle

# Done!

